---
title: "Computational Statistics-Report"
author: "David Niederkofler, Erlend Lokna"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(HDInterval)
library(R2OpenBUGS)
```



```{r}
mydata<-read.table("Report2_Dataset.txt", header=FALSE)
```
\section{Statistical Analysis of Covariates}
It is important to mention the use of notation before we proceed. We will in this section use the notation $\mathbf{x} = \{x_1, x_2, \cdots, x_n\}$ as our covariate vector.


\subsection{Ascicles}
```{r, include=FALSE}
mydata$V7<-as.numeric(mydata$V7)
```


\subsubsection{1.1 Model selection}
Since the Ascicles - covariate has a 0-1 outcome we can assume that it is Bernoulli distributed with parameter $\theta$. A natural conjugate prior for the Bernoulli distribution is the Beta distribution. The posterior beta distribution for the parameter is given by
$$Beta(\theta | a + \sum_{i = 1}^nx_i, b + n- \sum_{i + 1}^n x_i)$$
\subsubsection{1.2 Results}
The following results where found using the posterior beta distribution with a=1 and b=1 (Uniform distributed) for the ascicles data:
```{r, echo=FALSE}
n<-length(mydata$V6[!is.na(mydata$V7)])
s<-sum(mydata$V7)
mean<-(2+s)/(4+n)
mode<-(1+s)/(2+n)
cil<-qbeta(0.05/2, 1+s, 1 + n - s)
ciu<-qbeta(1-0.05/2,1+s, 1 + n - s)
cat("Posterior mean: ",mean)
cat("Posterior mode: ",mode)
cat("Centered 95% Confidence Interval: [",cil,",",ciu,"]")
```
With the following HPD interval:
```{r, echo = FALSE}
hdi(rbeta(1e5, 1+s, 1 + n - s))
```




\subsection{1. Sex}
```{r, include=FALSE}
mydata$V6<-as.numeric(mydata$V6)
```
The sex of the patients is encoded in a binary variable, where $0$ means \emph{male} and $1$ means \emph{female}.
\subsubsection{1.1 Model selection}
We assume a Bernoulli model $Ber(\theta)$ for the sex of the patient conditional on one parameter $\theta$, the probability of the patient to be female. The density function is given by
\begin{equation}
f(x|\theta)=\theta^x(1-\theta)^{1-x},
\end{equation}
where $x \in \{0,1\}$. As a prior distribution for $\theta$ we use the natural conjugate family of the Bernoulli distribution, namely the Beta distribution, $Beta(a,b)$, with two shape parameters $a=b=2$ to give more weight to the middle of the interval $[0,1]$, knowing how females and males are represented in the general population. The density is given by
\begin{equation}
h(\theta)=\frac{\Gamma(a+b)}{\Gamma(a) \Gamma(b)} \theta^{a-1} (1-\theta)^{b-1},
\end{equation}
for $\theta \in [0,1].$
\subsubsection{1.2 Results}
From the given dataset we get the sample size $n$ and the sum of the observations $s$:
```{r}
n<-length(mydata$V6[!is.na(mydata$V6)])
s<-sum(mydata$V6)
n
s
```
Therefore the posterior distribution is $Beta(2+s,2+n-s)$, which turns out to be $Beta(278,38)$.
From that we get
```{r, echo=FALSE}
mean<-(2+s)/(4+n)
mode<-(1+s)/(2+n)
cil<-qbeta(0.05/2,278,38)
ciu<-qbeta(1-0.05/2,278,38)
cat("Posterior mean: ",mean)
cat("Posterior mode: ",mode)
cat("Centered 95% Confidence Interval: [",cil,",",ciu,"]")
```
And the HPD confidence Interval calculates to:
```{r}
tst<-rbeta(1e5,278,38)
hdi(tst)
```
\subsection{2. Spiders}
```{r, include=FALSE}
mydata$V9<-as.numeric(mydata$V9)
```
The presence of spiders is encoded in a Binary variable, where $1$ means spiders are present.

\subsubsection{2.1 Model selection}
We assume a Bernoulli model $Ber(\theta)$ for the presence of spiders in patients conditional on one parameter $\theta$, the probability of the presence of spiders in the patient. The density function is given as stated earlier.
As a prior distribution for $\theta$ we use the natural conjugate family of the Bernoulli distribution, namely the Beta distribution, $Beta(a,b)$, with two shape parameters $a=b=1$, because we have no prior information. The density is given as above.

\subsubsection{2.2 Results}
From the given dataset we get the sample size $n$ and the sum of the observations $s$:
```{r}
n<-length(mydata$V9[!is.na(mydata$V9)])
s<-sum(mydata$V9)
n
s
```
Therefore the posterior distribution is $Beta(1+s,1+n-s)$, which turns out to be $Beta(91,223)$.
From that we get
```{r, echo=FALSE}
mean<-(2+s)/(4+n)
mode<-(1+s)/(2+n)
cil<-qbeta(0.05/2,91,223)
ciu<-qbeta(1-0.05/2,91,223)
cat("Posterior mean: ",mean)
cat("Posterior mode: ",mode)
cat("Centered 95% Confidence Interval: [",cil,",",ciu,"]")
```
And the HPD confidence interval calculates to:
```{r}
tst<-rbeta(1e5,91,223)
hdi(tst)
```
\subsection{3. Hepatomegaly}
```{r, include=FALSE}
mydata$V8<-as.numeric(mydata$V8)
```
The presence of hepatomegaly is encoded in a Binary variable, where $1$ means hepatomegaly is present.

\subsubsection{3.1 Model selection}
We assume a Bernoulli model $Ber(\theta)$ for the presence of hepatomegaly in the patient, conditional on one parameter $\theta$, the probability of the presence of hepatomegaly in the patient. The density function is given as stated earlier.
As a prior distribution for $\theta$ we use the natural conjugate family of the Bernoulli distribution, namely the Beta distribution, $Beta(a,b)$, with two shape parameters $a=b=1$, because we have no prior information. The density is given as above.

\subsubsection{3.2 Results}
From the given dataset we get the sample size $n$ and the sum of the observations $s$:
```{r}
n<-length(mydata$V8[!is.na(mydata$V8)])
s<-sum(mydata$V8)
n
s
```
Therefore the posterior distribution is $Beta(1+s,1+n-s)$, which turns out to be $Beta(161,153)$.
From that we get
```{r, echo=FALSE}
mean<-(2+s)/(4+n)
mode<-(1+s)/(2+n)
cil<-qbeta(0.05/2,161,153)
ciu<-qbeta(1-0.05/2,161,153)
cat("Posterior mean: ",mean)
cat("Posterior mode: ",mode)
cat("Centered 95% Confidence Interval: [",cil,",",ciu,"]")
```
And the HPD confidence interval calculates to:
```{r}
tst<-rbeta(1e5,161,153)
hdi(tst)
```
\subsection{4. Histologic stage}
The Histologic stage of the disease is a number in $\{1,2,3,4\}$, where the stage increases with severeness. We will give here the frequencies of the stages in the dataset.
```{r, echo=FALSE}
mydata$V20<- factor(mydata$V20,levels=c(1,2,3,4),labels=c("1","2","3","4"))
summary(mydata$V20)
```
We see that, most patients have been diagnosed in the last to stages of the disease.



\subsection{5. Age}

The age of the patient in days. 

\subsubsection{5.1 Model selection}
The data seems to follow a poisson distribution $Poi(\lambda)$. Using the non informative Jeffreys prior, we can derive that the posterior for the parameter $\lambda$ is Gamma distributed.

$$\theta | x \sim Gamma(\alpha=\frac{1}{2} + \sum_{i = 1}^n x_i, \beta = n)$$
\subsubsection{5.2 Results}



$$\theta | x \sim Gamma(\frac{1}{2} + s, n)$$

```{r, echo=FALSE}
poisson_jeffrey <- function(data){

  new_data = suppressWarnings(as.numeric(data))

  alpha = 1/2 + sum(new_data, na.rm=TRUE)
  beta = length(new_data)
  mean = alpha / beta
  variance = alpha / beta^2
  
  cat("posterior distribution:", "Gamma(",alpha,",",beta,")\n")
  cat("mean:", mean, "\n")
  cat("variance:", variance, "\n")
  
  cat("HPD intervall: \n")
  tst = rgamma(1e5, alpha, beta )
  hdi(tst)
}

poisson_jeffrey(mydata$V5)
```

\subsection{6. Cholesterol}


\subsubsection{6.1 Model selection}
We assume that the data is sampled from a poisson, $Poi(\lambda)$, distribution, and we use the non informative Jeffreys prior for the rate parameter in the bayesian analysis.

$$\mathbf{x} \sim Poi(\lambda)$$
$$h(\lambda) \propto \lambda^{-\frac{1}{2}}$$

\subsubsection{6.2 Results}
```{r, echo=FALSE}
poisson_jeffrey(mydata$V12)
```

\subsection{7. Urine}

\subsubsection{7.1 Model selection}
$$\mathbf{x} \sim Poi(\lambda)$$
$$h(\lambda) \propto \lambda^{-\frac{1}{2}}$$

\subsubsection{7.2 Results}
```{r, echo=FALSE}
poisson_jeffrey(mydata$V14)
```
\subsection{8 SGOT}

\subsubsection{8.1 Model selection}
$$\mathbf{x} \sim Poi(\lambda)$$
$$h(\lambda) \propto \lambda^{-\frac{1}{2}}$$
\subsubsection{8.2 Results}
```{r}
poisson_jeffrey(mydata$V16)
```

\subsection{9. Plateles}
\subsubsection{9.1 Model selection}
$$\mathbf{x} \sim Poi(\lambda)$$
$$h(\lambda) \propto \lambda^{-\frac{1}{2}}$$
\subsubsection{9.2 Results}
```{r}
poisson_jeffrey(mydata$V18)
```
\subsection{10. Prothrombin}
\subsubsection{10.1 Model selection}
$$\mathbf{x} \sim Poi(\lambda)$$
$$h(\lambda) \propto \lambda^{-\frac{1}{2}}$$
\subsubsection{10.2 Results}
```{r}
poisson_jeffrey(mydata$V19)
```


\subsection{11. Bilirubin}
The serum bilirubin of the patients is given in mg/dl.

\subsubsection{11.1 Model Selection}
We assume by inspecting the histogramm plot,

```{r, echo=FALSE}
bilirubin<-as.numeric(mydata$V11)
hist(bilirubin)

```

that the data follows a exponential distribution with parameter $\lambda$. Density is given by
\begin{equation}
f(x|\lambda)=\lambda e^{-\lambda x}
\end{equation}
As a prior for $\lambda$ we use, the jeffreys non-informative prior, namely: $h(\lambda) \propto \frac{1}{\lambda}$.

\subsubsection{11.2 Results}
From the data we get the number of samples $n$ and the sum of the samples $s$ as
```{r, echo=FALSE}
n<-length(bilirubin[!is.na(bilirubin)])
s<-sum(bilirubin)
n
s
```
That means the posterior distribution for $\lambda$ is $Gamma(n,s)$.
Which turns out to be $Gamma(312,1015.9)$. From that we get
```{r, echo=FALSE}
mean<-n/s
mode<-(n-1)/(s)
cil<-qgamma(0.05/2,312,1015.9)
ciu<-qgamma(1-0.05/2,312,1015.9)
cat("Posterior mean: ",mean)
cat("Posterior mode: ",mode)
cat("Centered 95% Confidence Interval: [",cil,",",ciu,"]")
```
And the HPD confidence interval calculates to:
```{r}
tst<-rgamma(1e5,312,1015.9)
hdi(tst)
```

\subsection{12. Albumin}
The Albumin is given in mg/dl.

\subsubsection{12.1 Model selection}
By the histogram plot of the data we see,
```{r echo=FALSE}
albumin<-as.numeric(mydata$V13)
hist(albumin)
```


that albumin could be gamma distributed with shape and rate parameters $a$ and $b$. We assume prior independence between $a$ and $b$ and use the marginal prior distributions $Gamma(0.001,0.001)$ for both of them.

\subsubsection{12.2 Results}
Using OpenBUGS and MCMC methods, we get posterior information about the parameters $a$ and $b$:
```{r}
n<-length(albumin[!is.na(albumin)])
X<-albumin
data1<-list("X","n")
params<-c("a" , "b")
inits<-list(a=1,b=1)
fit1<-bugs(data=data1,inits=list(inits),parameters.to.save=params,"model_albu.txt",n.chains=1, n.iter=20000, n.burnin = 10000,debug=FALSE, save.history = FALSE, DIC=TRUE)
fit1$summary
```

\subsection{13. Alkaline}
The data contains the units of alkaline phosphatase per liter of the patients. 

\subsubsection{13.1 Model selection}
Since the units of alkaline per liter are integers, we assume that it is a counting process. Therefore we want to assume, that the data is poisson distributed conditional on one parameter $\lambda$. The histogram plot justifies our assumptions:
```{r echo=FALSE}
alkaline<-as.numeric(mydata$V15)
hist(alkaline)
```
The density function of a single obervation is given as
\begin{equation}
f(x|\lambda)= e^{-\lambda} \frac{\lambda^x}{x!}
\end{equation}
As a prior for $\lambda$ we use the natural conjugate prior of the poisson distribution which is the gamma distribution. To not give a lot of prior information, we will use $Gamma(0.001,0.001)$.

\subsubsection{13.2 Results}
From our data we get the sample size $n$ and the sum $s$ over the sample:

```{r echo=FALSE}
n<-length(alkaline[!is.na(alkaline)])
s<-sum(alkaline)
n
s
```
The posterior distribution for $\lambda$ is given by $Gamma(s+0.001,n+0.001)$ which in our case results to $Gamma(618588.601,312.001)$. This yields to:
```{r, echo=FALSE}
mean<-(s+0.001)/(n+0.001)
mode<-(s+0.001-1)/(n+0.001)
cil<-qgamma(0.05/2,618588.601,312.001)
ciu<-qgamma(1-0.05/2,618588.601,312.001)
cat("Posterior mean: ",mean)
cat("Posterior mode: ",mode)
cat("Centered 95% Confidence Interval: [",cil,",",ciu,"]")
```
And the HPD confidence interval calculates to:
```{r}
tst<-rgamma(1e5,618588.601,312.001)
hdi(tst)
```

\subsection{14. Triglicerides}
Triglicerides of the patients in mg/dl.

\subsubsection{14.1 Model selection}
By the histogram plot of the data we see,
```{r echo=FALSE}
triglicerides<-as.numeric(mydata$V17)
hist(triglicerides)
```
that triglicerides could be gamma distributed with shape and rate parameters $a$ and $b$. We assume prior independence between $a$ and $b$ and use the marginal prior distributions $Gamma(0.001,0.001)$ for both of them.

\subsubsection{14.2 Results}

By OpenBUGS and MCMC methods we get posterior information about the parameters $a$ and $b$:
```{r}
n<-length(triglicerides[!is.na(triglicerides)])
X<-triglicerides
data1<-list("X","n")
params<-c("a" , "b")
inits<-list(a=1,b=1)
fit1<-bugs(data=data1,inits=list(inits),parameters.to.save=params,"model_albu.txt",n.chains=1, n.iter=20000, n.burnin = 10000,debug=FALSE, save.history = FALSE, DIC=TRUE)
fit1$summary
```


\section{Weibull Survival Analysis}
 We will use a survival model, based on a hazard function, conditional on regression parameters and dependent on (now assumed) deterministic covariates. The hazard function is given by
 \begin{equation}
 \lambda(t|\alpha, \beta, \delta)=\delta \alpha t^{\alpha-1}e^{\beta^Tz}
 \end{equation}
where $z$ is the covariate vector.
```{r, echo=FALSE}
mydata[mydata=="."]<-NA
mydata<-na.omit(mydata)
mydata$V3[mydata$V3==1]<-0
mydata$V3[mydata$V3==2]<-1
```



# Appendix
## Bernoulli/Beta
A natural conjugate prior for the Bernoulli distribution is the Beta distribution. 
$$f(x_i|\theta) = \theta^{x_i}(1-\theta)^{1-x_i}$$
$$L(\mathbf{x} | \theta) = \theta^{\sum_{i=1}^n x_ i}(1- \theta)^{n - \sum_{i=1}^n x_i } $$
$$h(\theta) = Beta(a, b)$$
We proceed by calculating the posterior distribution for $\theta$
$$h(\theta | \mathbf{x}) \propto L(\mathbf{x} | \theta) h(\theta) = \theta^{\sum_{i=1}^n x_ i}(1- \theta)^{n - \sum_{i=1}^n x_i } \frac{1}{B(a, b)} \theta^{a-1}(1-\theta)^{b-1} I(0 < \theta < 1)$$
$$\propto Beta(\theta | a + \sum_{i = 1}^nx_i, b + n- \sum_{i + 1}^n x_i) $$

## Poisson/Gamma
If our data $X_1, \cdot, X_n$ are iid Poisson($\lambda$) distributed then a gamma($\alpha$, $\beta$) prior on $\lambda$ is a conjugate prior.
The Likelyhood function is:
$$L(\lambda | \mathbf{x}) = \prod_{i=1}^n \frac{e^{-\lambda}\lambda^{\sum_{i=1}^n x_i}}{x_i !} = \frac{e^{-\lambda}\lambda^{\sum_{i=1}^n x_i}}{\prod_{i=1}^n x_i !}$$

Our gamma prior has the expression:
$$ h(\lambda) = \frac{\beta^{\alpha}}{\Gamma(\alpha)} \lambda^{\alpha-1} e^{-\beta \lambda} $$

Using bayes rule we find the following posterior:
$$h(\lambda | \mathbf{x}) \propto h(\lambda) L(\mathbf{x}|\lambda) \propto \lambda^{\sum_{i=1}^n x_i + \alpha -1} e^{-(n+\beta)\lambda}$$

$$\propto gamma(\sum_{i=1}^n x_i + \alpha, n+\beta)$$

## Poisson/Jeffreys prior

The density distribution for poisson is equal to
$$f(n|\lambda) = e^{-\lambda} \frac{\lambda^n}{n!}$$
The jeffreys prior $h(\theta$ is a non informative prior distrubution for a parameter space and its proportionality is expressed as

$$ h(\theta) \propto \sqrt{ det I(\theta)} $$
$$I(\theta) = -E[\frac{\partial^2}{\partial\theta^2} log f(x | \theta)] = \frac{1}{\theta}$$
And the following jeffreys prior is thus

$$h(\theta) \propto \theta^{-\frac{1}{2}} I_{\theta > 0}$$


The posterior is calculated as follows

$$h(\theta|x) \propto f(\mathbf{x} | \theta) h(\theta) \propto e^{-n\theta} \theta^{-\frac{1}{2} + \sum_{i = 1}^n x_i}$$ 
which is in fact a gamma distribution
$$\theta | x \sim Gamma(\alpha=\frac{1}{2} + \sum_{i = 1}^n x_i, \beta = n)$$










