---
title: "Computational Statistics-Report"
author: "David Niederkofler, Erlend Lokna"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
mydata<-read.table("Report2_Dataset.txt", header=FALSE)
```
\section{Statistical Analysis of Covariates}
It is important to mention the use of notation before we proceed. We will in this section use the notation $\mathbf{x} = \{x_1, x_2, \cdots, x_n\}$ as our covariate vector.


\subsection{Bernoulli - Beta, Ascicles}

Since the Ascicles - covariate has a 0-1 outcome we can assume that it is Bernoulli distributed with parameter $\theta$. A natural conjugate prior for the Bernoulli distribution is the Beta distribution. Therefore we have the following initial information:
$$f(x_i|\theta) = \theta^{x_i}(1-\theta)^{1-x_i}$$
$$L(\mathbf{x} | \theta) = \theta^{\sum_{i=1}^n x_ i}(1- \theta)^{n - \sum_{i=1}^n x_i } $$
$$h(\theta) = Beta(a, b)$$
We proceed by calculating the posterior distribution for $\theta$
$$h(\theta | \mathbf{x}) \propto L(\mathbf{x} | \theta) h(\theta) = \theta^{\sum_{i=1}^n x_ i}(1- \theta)^{n - \sum_{i=1}^n x_i } \frac{1}{B(a, b)} \theta^{a-1}(1-\theta)^{b-1} I(0 < \theta < 1)$$
$$\propto Beta(\theta | a + \sum_{i = 1}^nx_i, b + n- \sum_{i + 1}^n x_i) $$

\subsubsection{1.1 Model selection}

\subsubsection{1.2 Results}
\subsection{1. Sex}
```{r, include=FALSE}
mydata$V6<-as.numeric(mydata$V6)
```
The sex of the patients is encoded in a binary variable, where $0$ means \emph{male} and $1$ means \emph{female}.
\subsubsection{1.1 Model selection}
We assume a Bernoulli model $Ber(\theta)$ for the sex of the patient conditional on one parameter $\theta$, the probability of the patient to be female. The density function is given by
\begin{equation}
f(x|\theta)=\theta^x(1-\theta)^{1-x},
\end{equation}
where $x \in \{0,1\}$. As a prior distribution for $\theta$ we use the natural conjugate family of the Bernoulli distribution, namely the Beta distribution, $Beta(a,b)$, with two shape parameters $a=b=2$ to give more weight to the middle of the interval $[0,1]$, knowing how females and males are represented in the general population.
\subsubsection{1.2 Results}
From the given dataset we get the sample size $n$ and the sum of the observations $s$:
```{r}
n<-length(mydata$V6[!is.na(mydata$V6)])
s<-sum(mydata$V6)
n
s
```
Therefore the posterior distribution is $Beta(2+s,2+n-s)$, which turns out to be $Beta(278,38)$.
From that we get
```{r, echo=FALSE}
mean<-(2+s)/(4+n)
mode<-(1+s)/(2+n)
cil<-qbeta(0.05/2,278,38)
ciu<-qbeta(1-0.05/2,278,38)
cat("Posterior mean: ",mean)
cat("Posterior mode: ",mode)
cat("Posterior 95% Confidence Interval: [",cil,",",ciu,"]")
```

