---
title: "Computational Statistics-Report"
author: "David Niederkofler, Erlend Lokna"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



```{r}
mydata<-read.table("Report2_Dataset.txt", header=FALSE)

```
\section{Statistical Analysis of Covariates}

It is important to mention the use of notation before we proceed. We will in this section use the notation $\mathbf{x} = \{x_1, x_2, \cdots, x_n\}$ as our covariate vector.


\subsection{Bernoulli - Beta, Ascicles}

Since the Ascicles - covariate has a 0-1 outcome we can assume that it is Bernoulli distributed with parameter $\theta$. A natural conjugate prior for the Bernoulli distribution is the Beta distribution. Therefor we have the following initial information:
$$f(x_i|\theta) = \theta^{x_i}(1-\theta)^{1-x_i}$$
$$L(\mathbf{x} | \theta) = \theta^{\sum_{i=1}^n x_ i}(1- \theta)^{n - \sum_{i=1}^n x_i } $$
$$h(\theta) = Beta(a, b)$$
We proceed by calculating the posterior distribution for $\theta$
$$h(\theta | \mathbf{x}) \propto L(\mathbf{x} | \theta) h(\theta) = \theta^{\sum_{i=1}^n x_ i}(1- \theta)^{n - \sum_{i=1}^n x_i } \frac{1}{B(a, b)} \theta^{a-1}(1-\theta)^{b-1} I(0 < \theta < 1)$$
$$\propto Beta(\theta | a + \sum_{i = 1}^nx_i, b + n- \sum_{i + 1}^n x_i) $$

\subsubsection{1.1 Model selection}

\subsubsection{1.2 Results}

